### 2013-04-15 doc/final_report/introduction.tex
I dun added mah fixes for amanda and maria's notes

Since the advent of the 'Digital Age' and the ability of computers to copy and
reproduce information for a negligible cost, the amount of information
available to researchers has been increasing drastically.  B-C Bj\"{o}rk (2009)
estimates that approximately 1.4 Million papers were published in the year 2006
alone\cite{bjork2009}. Moreover, the growing popularity of Open Access
publishing (making papers available to readers for free online\cite{Suber2012})
across most scientific disciplines\cite{bjork2009}\cite{harnad2004comparing} is
providing scientists and researchers with an even larger volume of information
to be processed. As available information increases, relevant material becomes
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. The problem is even more vital for
General Practitioners. Goldacre (2008:97) points out that ``...there have been an
estimated 15 million medical academic articles published so far, and 5000
journals published every month... picking out what's relevant is a gargantuan
task."\cite{goldacre2008bad}.

@@

\section{Project Objectives}
\label{sec:objectives}
Partridge's main objective is to provide reading recommendations and
information retrieval for scientists and researchers. This should reduce the
amount of extraneous information that users have to read for themselves by
helping them to quickly find information specific to their interests.
Partridge will achieve this through the use of several existing techniques in
Natural Language Processing and utilisation of SAPIENTA as discussed above.

@@

\begin{itemize}
\item The system provides search within specific CoreSC regions of a given
paper. This would allow users to find the most relevant papers to them. For
example, a user may search for "mass spectrometer" within the methodology
section of a paper.
\item Partridge allows authors and rightsholders of scientific papers to
upload their papers to the system for inclusion into the search system. This
would give Partridge itself more utility and would make involved authors more
accessible.
\item The system provides filtering of papers based upon their
specific domain (i.e. is the paper primarily concerned with methodology within
an experiment in chemistry or is it about ethics in psychological studies?) and
their results, whether the paper yielded positive, negative or inconclusive
evidence for a hypothesis. 
\item Depending upon the time constraints of the
project, it was hoped that Partridge would also offer a user 'profiling' system
that provides recommendations for researchers based on their reading history.
This feature would help users find relevant papers more quickly or find
research that they may have otherwise overlooked.
\end{itemize}

@@

There are already a large number of information retrieval and recommendation
systems for scientific publications.  Many of these systems, such as AGRICOLA
(\url{http://agricola.nal.usda.gov/}), the Cochrane
Library(\url{http://www.thecochranelibrary.com/}) and
EconLit(\url{http://www.aeaweb.org/econlit/index.php}) index only publications
from predefined journals or topics (for the above examples, Agriculture,
Biology and Economics respectively). These sites can be very helpful for
researchers looking for literature covering specific topics. Filtering ever
smaller collections of academic literature allows for more precision; the
narrow subject area facilitating specialised search and indexing features that
help researchers locate relevant publications and articles even more quickly.
Figure \ref{fig:agricola} shows the extra search and filter controls offered by
the AGRICOLA repository which allow the user to filter papers by audience and
specific fields within Agriculture such as ``Animals and Livestock" and ``Food
and Nutrition".

@@

In contrast, there are also a number of interdisciplinary indexing system
systems such as arXiv( \url{http://arxiv.org/}), Mendeley
(\url{http://www.mendeley.com/research-papers/}) and JSTOR
(\url{http://www.jstor.org/}), that try to incorporate wide ranges of papers
from as many disciplines as possible. The traits of these systems often
complement those of their domain-specific counterparts; they provide a
comprehensive collection of literature but insufficient filtering and indexing
capabilities.

One of the most publicised and well known interdisciplinery scientific
literature search systems is Google Scholar (\url{http://scholar.google.com}).
As can be seen in Figure \ref{fig:scholar_basic}, This is an adaptation of
Google's general search engine to specifically index and search scientific
papers.  Google offers advanced query options specific to Scholar that allow
searching by author, year and for words that occur only in the document title
as shown in figure \ref{fig:scholar_advanced}. Whilst this does deal with the
problem of `information overload' by providing fine control over the
information returned from searches,  the user is still required to have a very
good idea of what they are looking for in terms of keywords and/or specific
authors. It is possible that a user would not know what they are looking for
until they've seen it. Even if the user has a set of keywords to search for,
they can only search the title of the paper or the content as a whole. This
means that users who want to find a particular phrase within a CoreSC part of
the paper (e.g.  only look for this phrase in the `Result' section of the
paper) are unable to get results at their desired level of detail.

@@

Search engines such as those discussed above are fairly useful for helping
people to find relevant literature and information when they have some idea of
what they are looking for. However, they demand precision in the use of
keywords and phrases which is unattainable by most researchers. Moreover, users
can get completely different result sets by entering synonyms. In comparison,
Recommender Systems may perform with more accuracy and better inclusion when
a search term cannot be clearly defined.

@@

Collaborative Recommender Services like Goodreads
(\url{http://www.goodreads.com/}) and CiteUlike
(\url{http://www.citeulike.org/}) allow users to register their interest in
specific authors and subjects. This allows the sites to build up a profile of
the sorts of materials that its users might be interested in and provide lists
of recommendations as in Figure \ref{fig:social_indexes}.

\begin{figure}[!hbt]
        \centering
        \begin{subfigure}[b]{0.50\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images/goodreads_index.png}
                \caption{Goodreads user profile page}
                \label{fig:goodreads_index}
        \end{subfigure}%
        \begin{subfigure}[b]{0.50\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images/citeulike_index.png}
                \caption{a CiteULike user profile page}
                \label{fig:citeulike_index}
        \end{subfigure}

        \caption{Goodreads and citeulike social Recommender systems}
        \label{fig:social_indexes}
\end{figure}

These systems have the ability to make recommendations to the user without
requiring specific keywords or search terms. They do this by learning the
user's profile and taking into account the preferences of their 'friends' and
their browsing history. However, the above-named systems do not take into
account the content of the paper or book. They only deal with metadata as can
be seen in Figure \ref{fig:social_searches}. This means that important
discriminatory information that could be contained within the actual document
content is overlooked completely.

\begin{figure}[!hbt]
        \centering
        \begin{subfigure}[b]{0.50\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images/goodreads_search.png}
                \caption{Goodreads advanced search page}
                \label{fig:goodreads_search}
        \end{subfigure}%
        \begin{subfigure}[b]{0.50\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images/citeulike_search.png}
                \caption{CiteULike advanced search options}
                \label{fig:citeulike_search}
        \end{subfigure}

        \caption{Goodreads and citeulike search only deal with metadata.}
        \label{fig:social_searches}
\end{figure}

@@

Hybrid Recommender Systems are designed to combine the Content-based and
Collaboration-based models, employing aspects of both techniques. Huang
\emph{et al.}(2002) describe a graph-based approach to combine content and
collaborative data from books, customers and transactions taken from an online
bookseller\cite{Huang:2002:GRS:544220.544231}. Combining both sources of
information boosted the effectiveness of the recommendation engine and
\emph{'...the system gained improvement with respect to both precision and
recall...' (Ibid.)} A scientific literature recommender system could make use
of a similar hybrid technique in order to provide effective recommendations to
researchers.

@@

Human literature is generally written in human languages for the consumption of
other humans. Chomsky (1956) defines a hierarchy for the classification of
formal grammars that describe languages \cite{chomsky1956three}. Although it
can be debated where exactly on the hierarchy human language fits (
Reich(1969)\cite{reich1969finiteness} argues that Natural Languages can be
considered finite state and Matthews(1979)\cite{matthews1979grammatical} argues
that Natural Languages cannot even be considered recursively enumerable, cited
in Kornai(1985)\cite{kornai1985}) it is clear that they are not in the same
category as most forms of machine language. Literature written for humans must
be interpretted for machines and this is where Natural Language Parsing is
useful.

@@

\subsection{Natural Language Processing in Scientific Papers}

Liakata \emph{et al.} (2012) describe a system for automatically processing and
classifying sentences in a research paper according to the scientific concept
they describe (e.g. a sentence can be categorised as a hypothesis, background
information, a method or similar)\cite{citeulike:10444769}. SAPIENTA
(\url{http://www.sapientaproject.com}) is a machine learning application,
trained using a corpus of physical chemistry and biochemistry research papers
that were manually annotated using the Core Information about Scientific
Papers (CISP)\cite{LIAKATA10.644} schema.

CISP, Soldatova and Liakata (2007), is a way to formally represent core
scientific concepts (CoreSC), e.g. background, hypothesis, method etc., that
should be present in the articles in a logical
ontology\cite{soldatova2007ontology}. Subsequent work implements this set of
concepts as a three layered sentence based annotation scheme (CoreSC scheme)
where annotations are encoded in Extended Markup Language
(XML)\cite{LIAKATA10.644}. An intelligent academic information retrieval system
could utilise this information in order to provide better filtering and search
capabilities for researchers.

>>>>

Since the advent of the `Digital Age' and the ability of computers to copy and
reproduce information for a negligible cost, the amount of information
available to researchers has been increasing drastically.  B-C Bj\"{o}rk (2009)
estimates that approximately 1.4 Million papers were published in the year 2006
alone\cite{bjork2009}. Moreover, the growing popularity of Open Access
publishing (making papers available to readers for free online\cite{Suber2012})
across most scientific disciplines\cite{bjork2009}\cite{harnad2004comparing} is
providing scientists and researchers with an even larger volume of information
to be processed. As available information increases, relevant material becomes
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. The problem is even more vital for
General Practitioners. Goldacre (2008:97) points out that ``...there have been an
estimated 15 million medical academic articles published so far, and 5000
journals published every month... picking out what's relevant is a gargantuan
task."\cite{goldacre2008bad}.

@@

\section{Project Objectives}
\label{sec:objectives}
Partridge's main objective is to provide reading recommendations and
information retrieval for scientists and researchers. This should reduce the
amount of extraneous information that users have to read for themselves by
helping them to quickly find information specific to their interests.
Partridge will achieve this through the use of several existing techniques in
Natural Language Processing.

@@

\begin{itemize}
\item The system provides search within specific sentence-based zones of a given
paper. This would allow users to find the most relevant papers to them. For
example, a user may search for ``mass spectrometer" within the methodology
section of a paper.
\item Partridge allows authors and rightsholders of scientific papers to
upload their papers to the system for inclusion into the search system. This
would give Partridge itself more utility and would make the works of involved authors more
accessible to researchers.
\item The system provides filtering of papers based upon their type (i.e Review
Papers, Research Papers and Case Studies). It also was hoped that Partridge would
also provide filtering by subject domain (i.e. is the paper primarily concerned with methodology within
an experiment in chemistry or is it about ethics in psychological studies?) and
their results, whether the paper yielded positive, negative or inconclusive
evidence for a hypothesis. 
\item A `paper profile' page is generated for each paper included in Partridge
which provides graphical representation of some of the analysis carried out on
the paper and helps the user to begin to judge the paper's content.
\item Depending upon the time constraints of the
project, it was hoped that Partridge would also offer a user `profiling' system
that provides recommendations for researchers based on their reading history.
This feature would help users find relevant papers more quickly or find
research that they may have otherwise overlooked.
\end{itemize}

@@

There are already a large number of information retrieval and recommendation
systems for scientific publications.  Many of these systems, such as AGRICOLA
(\url{http://agricola.nal.usda.gov/}), the Cochrane
Library(\url{http://www.thecochranelibrary.com/}) and
Textpresso (\url{http://www.textpresso.org/}) index only publications
from predefined journals or topics (for the above examples, Agriculture,
Biology and Bioinformatics respectively). These sites can be very helpful for
researchers looking for literature covering specific topics. Filtering ever
smaller collections of academic literature allows for more precision; the
narrow subject area facilitating specialised search and indexing features that
help researchers locate relevant publications and articles even more quickly.
Figure \ref{fig:agricola} shows the extra search and filter controls offered by
the AGRICOLA repository which allow the user to filter papers by audience and
specific fields within Agriculture such as ``Animals and Livestock" and ``Food
and Nutrition".

@@

In contrast, there are also a number of interdisciplinary indexing systems and
online journals such as arXiv( \url{http://arxiv.org/}),
PloSOne(\url{http://plosone.org/}), and JSTOR (\url{http://www.jstor.org/}),
that try to incorporate wide ranges of papers from as many disciplines as
possible. The traits of these systems often complement those of their
domain-specific counterparts; they provide a comprehensive collection of
literature but insufficient filtering and indexing capabilities.

One of the most publicised and well known interdisciplinery scientific
literature search systems is Google Scholar (\url{http://scholar.google.com}).
As can be seen in Figure \ref{fig:scholar_basic}, This is an adaptation of
Google's general search engine to specifically index and search scientific
papers.  Google offers advanced query options specific to Scholar that allow
searching by author, year and for words that occur only in the document title
as shown in figure \ref{fig:scholar_advanced}. Whilst this does deal with the
problem of `information overload' by providing fine control over the
information returned from searches,  the user is still required to have a very
good idea of what they are looking for in terms of keywords and/or specific
authors. It is possible that a user would not know what they are looking for
until they've seen it. Even if the user has a set of keywords to search for,
they can only search the title of the paper or the content as a whole. This
means that users who want to find a particular phrase within a specific
subsection of the paper (e.g.  only look for this phrase in the `Result'
section of the paper) are unable to get results at their desired level of
detail.

@@

Search engines such as those discussed above are fairly useful for helping
people to find relevant literature and information when they have some idea of
what they are looking for. However, they demand precision in the use of
keywords and phrases, often leaving researchers looking for literature on highly
specific topics frustrated. Moreover, users can get completely different result
sets by entering synonyms. In comparison, Recommender Systems may perform with
more accuracy and better inclusion when a search term cannot be clearly
defined.

@@

Collaborative Recommender Services like Mendeley
(\url{http://www.mendeley.com/}), ResearchGate
(\url{http://www.researchgate.net/})  and CiteUlike
(\url{http://www.citeulike.org/}) allow users to register their interest in
specific authors and subjects. This allows the sites to build up a profile of
the sorts of materials that its users might be interested in and provide lists
of recommendations.

These systems have the ability to make recommendations to the user without
requiring specific keywords or search terms. They do this by learning the
user's profile and taking into account the preferences of their `friends' and
their browsing history. However, the above-named systems do not take into
account the content of the paper or book. They only deal with the relationships
between users and their reading history. This means that important
discriminatory information that could be contained within the actual document
content is overlooked completely.

@@

Hybrid Recommender Systems are designed to combine the Content-based and
Collaboration-based models, employing aspects of both techniques. Huang
\emph{et al.}(2002) describe a graph-based approach to combine content and
collaborative data from books, customers and transactions taken from an online
bookseller\cite{Huang:2002:GRS:544220.544231}. Combining both sources of
information boosted the effectiveness of the recommendation engine and
the system gained improvement with respect to both precision and
recall \emph{(Ibid.).} A scientific literature recommender system could make use
of a similar hybrid technique in order to provide effective recommendations to
researchers.

@@

Human literature is generally written in human languages for the consumption of
other humans. Chomsky (1956) defines a hierarchy for the classification of
formal grammars that describe languages \cite{chomsky1956three}. Although it
can be debated where exactly on the hierarchy human language fits (
Reich(1969)\cite{reich1969finiteness} argues that Natural Languages can be
considered finite state and Matthews(1979)\cite{matthews1979grammatical} argues
that Natural Languages cannot even be considered recursively enumerable, cited
in Kornai(1985)\cite{kornai1985natural}) it is clear that they are not in the same
category as most forms of machine language. Literature written for humans must
be interpretted for machines and this is where Natural Language Parsing is
useful.

@@

\subsection{Natural Language Processing in Scientific Papers}
\label{sec:nlpsapienta}

Liakata \emph{et al.} (2012) describe a system for automatically processing and
classifying sentences in a research paper according to the scientific concept
they describe (e.g. a sentence can be categorised as a hypothesis, background
information, a method or similar)\cite{citeulike:10444769}. SAPIENTA
(\url{http://www.sapientaproject.com}) is a machine learning application,
trained using a corpus of physical chemistry and biochemistry research papers
that were manually annotated using the Core Scientific Concepts
(CoreSC)\cite{LIAKATA10.644} scheme.

CoreSC Liakata \emph{et al.} (2010), derived from earlier work in which Core
Information about Scientific Papers (CISP) is formalised as a way of improving
better metadata for papers\cite{soldatova2007ontology}, is a way to formally
represent core scientific concepts (CoreSC), e.g. background, hypothesis,
method etc., that should be present in scientific articles in a logical
ontology\cite{LIAKATA10.644}. It is implemented as a three layered sentence
based annotation scheme  where annotations are encoded in Extended Markup
Language (XML). CoreSC labels can be allocated to all sentences in a
scientific paper in order to identify which scientific concept each sentence
encapsulates. For example, a sentence with the CoreSC label ``Hypothesis"
conveys some expectation of an experiment or action. An example ``Hypothesis"
identified in a Bioinformatics paper used for testing is:

\begin{quotation}
We assume that, as in the HACA process, condensation is initiated by radical
formation, and then consider reactions of phenyl, naphthyl and anthracenyl with
another PAH to yield for example fluoranthene or
perylene\cite{unterreiner2004reaction}.
\end{quotation}

Alternatively, a ``Background'' CoreSC sentence may refer to prior work or
merely declare a point of interest to the reader. An example from the same
paper is:

\begin{quotation}
The central issue is here the formation of polycyclic aromatic hydrocarbons
(PAHs), which are the main constituents of soot; the PAHs are also health
hazards since they are carcinogenic\cite{unterreiner2004reaction}.
\end{quotation}

An intelligent academic information retrieval system could utilise this
information in order to provide better filtering and search capabilities for
researchers. The ability to search for phrases and keywords by CoreSC would
facilitate context-aware keyword search, that allows researchers to only accept
papers where a term appears in sentences with a specific CoreSC label. This
could be used to greatly improve both the precision with which researchers are
able to perform searches for scientific literature and the accuracy of those
searches in terms of relevance to the reader.

### 2013-04-15 doc/final_report/methodology.tex
I dun added mah fixes for amanda and maria's notes

Having analysed some of the existing development methodologies and determined
that most of the common methodologies were not suitable for Partridge. A custom
methodology was formulated and support tools chosen. Task analysis was then
performed and a project timeline generated. With a timeline in place, the
initial design phase for Partridge and its subsystems could begin.

>>>>

Having analysed some of the existing development methodologies and determined
that most of the common methodologies were not suitable for Partridge, a custom
methodology was formulated and support tools chosen. Task analysis was then
performed and a project timeline generated. With a timeline in place, the
initial design phase for Partridge and its subsystems could begin.

### 2013-04-15 doc/final_report/report.tex
Added some images and final changes
I dun added mah fixes for amanda and maria's notes

A preprocessing system system that exploits existing tools such as SAPIENTA and
PDFX was used to analyse a large sample of research papers, labelling the
presence of core scientific concepts (CoreSCs) within each paper and extracting
metadata. Decision trees were also trained to classify papers into categories
such as `Research Paper', `Review Paper' and `Case Study' based upon the
relative proportions of core scientific concepts in the papers.

A web interface was built to facilitate user search of papers based upon the
contents of specific CoreSCs. Filtering of papers using their type as
determined by the preprocessor was also built in and a page allowing authors and
rightsholders to submit their own papers for analysis was created to allow
the scientific community to extend the utility of the system.

@@

\chapter{Evaluation}
\input{evaluation}

>>>>

A preprocessing system that exploits existing Natural Language Processing tools
was used to analyse a large sample of research papers, labelling the presence
of conceptual zones within each paper and extracting metadata.  Decision trees
were also trained to classify papers into well established type categories
based upon the relative proportions of the types of conceptual zones
present within the papers.

A web interface was built to facilitate fine grain search, making use of the
zone information captured during preprocessing. Filtering of papers using their
type as determined by the preprocessor was also built in. Individual papers
returned during searches are presented in their own profile page that provides
charts and extra metadata to help researchers to understand the paper's
content. A page allowing authors and rightsholders to submit their own papers
for analysis was also created to allow the scientific community to extend the
impact and utility of the system.

The machine learning systems used within the project have been evaluated and
found to be accurate to 87.9\% when classifying a paper's type. The possibility
of extending the machine learning capabilities of the system to categorise
papers by subject domain and outcome of experimental results is discussed as
well as the potential for further improvements to the user interface.

A working version of the software has been made generally available at
\url{http://farnsworth.papro.org.uk} and can be explored and used by the
general public. It is hoped that over the coming months, the system will grow
to become a far reaching, widely used tool for academic research.

@@

\chapter{Critical Evaluation}
\input{evaluation}

### 2013-04-15 doc/final_report/testing.tex
I dun added mah fixes for amanda and maria's notes

Receiver Operating Characteristics (ROC) Charts are used to visually represent
the trade-off between True Positives(TP) and False Positives(FP) as identified
by the classifier under evaluation. Witten (2005) explains that ROC is a method
reappropriated from signal processing analysts who use it to visualise the ``...hit
rate and false alarm rate over noisy communication
channels\cite{witten2005data}." The ratio of TP to FP classifications is
measured with increasing numbers of test samples from left to right.

@@

These results show that Partridge's decision tree is very effective at
classifying the three document types in its database. It is also clear, however,
that whilst the classifier is outstanding at recognising research papers, it
sometimes struggles to discriminate between case study and review papers. This
may be because case studies can be considered a specific type of review paper
and the two classes share more similar CoreSC features than they do with
research papers.

@@

\section{ User Testing }

Raymond (1999) suggests that for developers of open source projects, ``...treating
your users as co-developers is your least-hassle route to rapid code
improvement and effective debugging\cite{raymond1999cathedral}." Partridge
hasn't been developed using an open methodology since, as a dissertation
project, it has been necessary to avoid complications around work ownership.
However, Raymond's notion that software should be released early and often and
tested on its user-base is an effective one and has been widely adopted by both
open and closed source developers\cite{linux2013}\cite{unity2013}.

>>>>

Receiver Operating Characteristics (ROC) Charts are used to visually represent
the trade-off between True Positives(TP); items that are correctly identified
as a given class and False Positives(FP); items that are identified as a given
class but are actually of aother class as identified
by the classifier under evaluation. Witten (2005) explains that ROC is a method
reappropriated from signal processing analysts who use it to visualise the ``...hit
rate and false alarm rate over noisy communication
channels\cite{witten2005data}." The ratio of TP to FP classifications is
measured with increasing numbers of test samples from left to right.

@@

These results show that Partridge's decision tree is very effective at
classifying the three document types in its database. It is also clear,
however, that whilst the classifier is outstanding at recognising research
papers, it sometimes struggles to discriminate between case study and review
papers. This may be because case studies share more similar CoreSC features
with review papers than they do with research papers.

@@

\section{ User Testing } 
\label{sec:user_testing}
Raymond (1999) suggests that for developers of open source projects, ``...treating
your users as co-developers is your least-hassle route to rapid code
improvement and effective debugging\cite{raymond1999cathedral}." Partridge
hasn't been developed using an open methodology since, as a dissertation
project, it has been necessary to avoid complications around work ownership.
However, Raymond's notion that software should be released early and often and
tested on its user-base is an effective one and has been widely adopted by both
open and closed source developers\cite{linux2013}\cite{unity2013}.
